Ideas

### 2023-07-16 Notes.tex
minima

We want a notion of shift-equivalence which ensures some properties
which are indispensable to allow to interpret derivations up-to shift
as equivalent concurrent computations which differ only for the fact
that independent steps have been moved around. When moving around steps, we should not loose track of the identity of the steps: this is essential to be able to have an event structure semantics with a clearly defined notion of event.

@@

\textbf{Question.} Taking the standard shift, with generic interchange pair, a a counterexample to property 1, is the one mentioned above.  Do we have a counterexample to property 2? Not clear to me.

@@

  \textbf{Questions}: Can we reduce the set of axioms? Do we need something else relating abstraction equivalence (iso of derivation) and shift? Relation with symmetries?

@@

\item Does the theory above apply to other cases. E.g., if one wants a
  proper notion of causality (avoiding or-causality) a possibility
  seems to be to restrict the applicability condition for rules, to
  prevent fusions to act on already merged items (see section V.D in
  UnstableFull). By the way, note that the satisfaction of the conditions
  on shift depends not only on the notion of shift, but also on the
  notion of rewriting. In this case we expect to have a prime event
  structure (bit this has to do with the second part of the work).

@@

Ideally, one would like to be able to work with an abstract notion of
shift-equivalence and just using its properties as singled out in the
previous part derive a weak prime domain or connected event structure
for it. This is probably difficult, but one might need to add some
futher properties to be proved for the specific notion of shift at
hand.

For this part, relevant material is section V of UnstableFull. We use abstraction equivalence (iso of derivations) $\equiv^{a}$ (def. 36), shift-equivalence $\shifteq_\sigma$ (Def. 38, traking the permutation) and their join $\equiv^c_\sigma$.

For linear rules on graphs it is known that two derivations are
related by $\rho \equiv^c_\sigma \rho'$ iff $\sigma$ is a consistent
permutation (def. 40, UnstableFull). Rough;y the idea is to take the colimit
and consider the rewriting steps over the colimit to deduce
dependencies. This was used in many places in the proofs, but for
left-linear rules the proof does not go through (only the implication
``if $\rho \equiv^c_\sigma \rho'$ then $\sigma$ is a consistent
permutation'' (Lemma 17(1)) seems to hold as the original proof do not
to rely on linearity, this should be double-checked, but the converse
is false). In general, the hope is to get rid of using consistent
permutation and use instead more asbtract properties of
shift-equivalence as discussed in the first part.

@@

\item
  \textbf{Lemma 19}.
  (1) says that if $\psi; \psi_1 \equiv^c_\sigma \psi'; \psi_1'$ and the permutation $\sigma$ sends $n$ steps of $\psi_1$ to $\psi'$ then such steps are not in $\psi$ and they will need to be in every $\phi_2$ such that $\psi; \phi_2 \equiv^c_\sigma \psi'; \phi_2'$. In particular there is $\phi_2$ consisting of exactly $n$ steps which allows for commutation. This allows essentially to deduce existence of minimal upper bounds, al with the same length. 
  Then (2) shows that all minimal upper bounds are isomorphic, so that we have a sup.

@@

  Here I think that the proofs are draft modifications of the original ones, changed after discovering the property on consistent permutation to be wrong. But they are drafty, to be redone maybe keeping some ideas.

>>>>

We want a notion of shift-equivalence that ensures some properties
which are indispensable to allow to interpret derivations up-to shift
as equivalent concurrent computations which differ only for the fact
that independent steps have been moved around. When moving around steps, we should not lose track of the identity of the steps: this is essential to be able to have an event structure semantics with a clearly defined notion of event.

@@

\textbf{Question.} Taking the standard shift with generic interchange pair, a counterexample to Property 1 is the one mentioned above.  Do we have a counterexample to Property 2? Not clear to me.

@@

  \textbf{Questions}: Can we reduce the set of axioms? Do we need something else relating abstraction equivalence (iso of derivations) and shift? Relation with symmetries?

@@

\item Does the theory above apply to other cases. E.g., if one wants a
  proper notion of causality (avoiding or-causality) a possibility
  seems to be to restrict the applicability condition for rules, to
  prevent fusions to act on already merged items (see section V.D in
  UnstableFull). By the way, note that the satisfaction of the conditions
  on shift depends not only on the notion of shift, but also on the
  notion of rewriting. In this case we expect to have a prime event
  structure (but this has to do with the second part of the work).

@@

Ideally, one would like to be able to work with an abstract notion of
shift-equivalence and just using its properties as singled out in the
previous part to derive a weak prime domain or connected event structure
for it. This is probably difficult, but one might need to add some
futher properties to be proved for the specific notion of shift at
hand.

For this part, relevant material is section V of UnstableFull. We use abstraction equivalence (iso of derivations) $\equiv^{a}$ (Def. 36), shift-equivalence $\shifteq_\sigma$ (Def. 38, traking the permutation) and their join $\equiv^c_\sigma$.

For linear rules on graphs it is known that two derivations are
related by $\rho \equiv^c_\sigma \rho'$ iff $\sigma$ is a consistent
permutation (Def. 40, UnstableFull). Roughly the idea is to take the colimit
and consider the rewriting steps over the colimit to deduce
dependencies. This was used in many places in the proofs, but for
left-linear rules the proof does not go through (only the implication
``if $\rho \equiv^c_\sigma \rho'$ then $\sigma$ is a consistent
permutation'' (Lemma 17(1)) seems to hold as the original proof do not
to rely on linearity, this should be double-checked, but the converse
is false). In general, the hope is to get rid of using consistent
permutations and use instead more abstract properties of
shift-equivalence as discussed in the first part.

@@

\item
  \textbf{Lemma 19}.
  (1) says that if $\psi; \psi_1 \equiv^c_\sigma \psi'; \psi_1'$ and the permutation $\sigma$ sends $n$ steps of $\psi_1$ to $\psi'$ then such steps are not in $\psi$ and they will need to be in every $\phi_2$ such that $\psi; \phi_2 \equiv^c_\sigma \psi'; \phi_2'$. In particular there is $\phi_2$ consisting of exactly $n$ steps which allows for commutation. This allows essentially to deduce existence of minimal upper bounds, all with the same length. 
  Then (2) shows that all minimal upper bounds are isomorphic, so that we have a sup.

@@

  Here I think that the proofs are draft modifications of the original ones, changed after discovering the property on consistent permutations to be wrong. But they are drafty, to be redone maybe keeping some ideas.
